\capitulo{1}{Introduction}

% Description of the work, the structure of the memory and the rest of the material delivered.

\nonzeroparskip Social networks are currently a fundamental aspect of society. People usually use social networks to share experiences, opinions, and aspects of their lives and interact with other people. Using social networks as a data source we can access a huge amount of information and be able to build accurate analyses on practically any topic.

\nonzeroparskip On the other hand, another aspect that has gradually permeated our society is the concept of subscriptions to services, be it music, movies, games, or almost any concept that we can think of. Not so many years ago, the concept of paying for subscriptions to services, where you do not actually own the content you pay for and instead get temporary access whose duration is defined by how long you continue to pay for the subscription, was relegated to very specific services and was not nearly as globalized as it is today.

\nonzeroparskip The global acceptance of subscription as a service is reflected in social networks, where users can comment on the different music, movie, and game platforms, turning each new release into a social phenomenon. In order to take advantage of both worlds, this project uses the social network Twitter to obtain information about the latest music listened to by users (by searching for a particular hashtag) and then consult the data of the song and the artist involved that Spotify, a platform based on music as a service, has.

\nonzeroparskip The development of the project has followed an agile methodology with two-week sprints, and has focused on learning and using recognized tools in the field of Big Data, such as Apache Airflow, Apache Spark, Apache Cassandra, Docker Compose and Flask, among others.